---
title: "Attrition - Modeling"
author: "Sterling Beason"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(class)
library(caret)
library(e1071)
library(tidyverse)
```

```{r}
dataRaw = read.csv('../data/CaseStudy2-data.csv', header = TRUE)

# Non-informative columns: ID, EmployeeNumber
# Useless columns with one value: EmployeeCount, Over18, StandardHours
data = dataRaw %>% select(-c('ID', 'EmployeeNumber', 'EmployeeCount', 'Over18', 'StandardHours'))
```

## Data Partitioning (Train/Test)
```{r}
# split data into training and testing
trainInd = createDataPartition(data$Attrition, times = 1, p = 0.75, list = FALSE)

train = data[trainInd,]
test = data[-trainInd,]

dim(train)
dim(test)
```

### Naive Bayes  - All Predictors
```{r}
model <- naiveBayes(Attrition ~ ., data = train)

pred <- predict(model, test)

# confusion Matrix
confusionMatrix(table(pred, test$Attrition))

```

### ~~Naive Bayes - Top Three (numeric) Predictors~~
Note: Top three of all predictors was resulting in zero-frequency issue. (OverTime + YearsWithCurrManager + StockOptionLevel) EDIT: so are the numeric...

Top three predictors identified in EDA using the MARS algorithm.
```{r}
modelReduced <- naiveBayes(Attrition ~ TotalWorkingYears + NumCompaniesWorked + YearsSinceLastPromotion, data = train)

predReduced <- predict(modelReduced, test)

confusionMatrix(table(predReduced, test$Attrition))

```

#### Lets try this with kNN
##### Identify best k
```{r, eval = FALSE}
set.seed(1432)
iterations = 250
numks = 60
splitPerc = .8

masterAcc = matrix(nrow = iterations, ncol = numks)

dataKnn <- data %>% select('Attrition', 'TotalWorkingYears', 'NumCompaniesWorked', 'YearsSinceLastPromotion')

for(j in 1:iterations)
{
  trainIndices = createDataPartition(dataKnn$Attrition, times = 1, p = splitPerc, list = FALSE)
  trainKnn = dataKnn[trainIndices,]
  testKnn = dataKnn[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(trainKnn[,2:4],testKnn[,2:4],trainKnn$Attrition, prob = TRUE, k = i)
    table(classifications,testKnn$Attrition)
    CM = confusionMatrix(table(classifications,testKnn$Attrition))
    masterAcc[j,i] = CM$overall[1]
  }
  
}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")

which.max(MeanAcc)
max(MeanAcc)

```

##### Run kNN with best K
```{r}
trainReduced <- train %>% select('Attrition', 'TotalWorkingYears', 'NumCompaniesWorked', 'YearsSinceLastPromotion')
testReduced <- test %>% select('Attrition', 'TotalWorkingYears', 'NumCompaniesWorked', 'YearsSinceLastPromotion')

knn <- knn(trainReduced[,2:4], testReduced[,2:4], trainReduced$Attrition, prob = TRUE, k = 8)

confusionMatrix(table(knn, testReduced$Attrition))
```

```{r}
trainNum <- train %>% select_if(is.numeric)
testNum <- test %>% select_if(is.numeric)

trainNum$Attrition <- train$Attrition
testNum$Attrition <- test$Attrition

names(trainNum)

knn <- knn(trainNum[,1:23], testNum[,1:23], trainNum$Attrition, prob = TRUE, k = 8)

confusionMatrix(table(knn, testNum$Attrition))
```